\textbf{Betelwise1: A Hybrid Conventional-CNN Approach for Efficient Planetary Seismic Event Detection and Analysis}

### 2025-05-24 paper.tex
update

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

@@

\section{Framework}
    %Overview
    Overview
    \subsection{Datasets and Input Seismogram Preparation}
    hello
        \subsubsection{Data Sources (Lunar and Martian)}
            hello

>>>>

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{enumitem}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

@@

\section{Framework}
This framework consists of two core components: an Initial Processing and Candidate
Selection (IPCS) stage leveraging conventional seismic algorithms, followed by a Machine Learning-based Event
Verification (MLEV) stage employing a Convolutional Neural Network (CNN).

The IPCS stage is responsible for initial signal conditioning and candidate event detection from raw seismograms. This
involves a sequence of operations including adaptive bandpass filtering to enhance relevant frequency content, outlier
removal to mitigate transient noise, data normalization for consistent scaling, and STA/LTA analysis to identify
potential seismic onsets. A key feature of this stage is a subsequent refinement step using characteristic function
slope analysis, designed to reduce false positives from the STA/LTA triggers by distinguishing between seismic event
profiles and noise bursts.

Candidate events identified and refined by the IPCS stage are then passed to the MLEV stage. For this stage, rather than
converting waveforms into 2D time-frequency representations, our framework utilizes 1D seismogram segments. These
segments are critically augmented with some auxiliary statistical features, such as standard deviations computed around
the potential arrival times, to provide the CNN with extra contextual information. A lightweight CNN then performs the
final classification, discerning true seismic events from the residual noise and false triggers that have passed the
conventional filtering. This hybrid architecture aims to optimize both computational efficiency and detection accuracy
for planetary missions.

    \subsection{Initial Processing and Candidate Events Selection(IPCS)}%TODO refine heading
    (This entire subsection is a major differentiator from purely ML-driven approaches like Civilini et al.'s direct
    spectrogram-to-CNN method. You're showing a sophisticated "front-end".)

        \subsubsection{Adaptive Bandpass Filtering}
        (Your method for "Freq AnimationFinding the Best Bandpass Frequency": power spectrum analysis or standard
            deviation analysis. Tunable parameters involved.)

            \subsubsection{Noise and Outlier Mitigation}
            (Techniques for removing short, high-amplitude noises and other artifacts.)

        \subsubsection{Data Normalization}
            (Method used for consistent scaling.)

        \subsubsection{STA\slash LTA Analysis for Initial Detection}
            (Standard STA\slash LTA application: window lengths, thresholds. This is a common baseline.)

        \subsubsection{Candidate Pick Refinement using Characteristic Function Analysis}
        \label{sec:candidate_refinement}
            (Detail this specific filtering mechanism based on rise/fall slopes. This is likely a novel or specifically
            adapted part of your conventional pipeline and should be emphasized.)

    \subsection{Machine Learning Stage: CNN-based Event Refinement}

    \subsubsection{Input Feature Construction for the CNN}
        (How candidate events from \ref{sec:candidate_refinement} are prepared for the CNN.) 1D Seismogram Segments:
        Length (e.g., 5565 samples), channels used. Auxiliary Features: Explicitly detail the "Standard deviation values
        before and after the Arrival time." Justify why these features are chosen and what information they provide to
        the network. (Highlighting point: Civilini et al. used 2D spectrograms. Your use of 1D waveforms + explicitly
        engineered auxiliary features is a key methodological difference.)

        \subsubsection{CNN Architecture}
        (Describe your "lightweight" CNN: number and type of layers (convolutional, pooling, fully connected),
        activation functions, filter sizes, strides, dropout, etc. Justify design choices if aiming for on-board
        efficiency.)

        \subsubsection{CNN Training and Validation}
        (Training data labeling (event vs. noise, P/S arrivals if applicable). Loss function. Optimizer. Learning rate.
        Batch size. Number of epochs. Validation strategy.) (Mention of using both lunar and Martian data for training,
        which is distinct from single-source transfer learning.)

        \subsubsection{Event Classification and Output}
        (How the CNN output (e.g., probability score) is used to make the final event/noise classification.
        Thresholding.)

        \subsection{Tunability and Adaptability of the Framework}
        (Briefly reiterate the tunable parameters (e.g., the six mentioned in your abstract) and how they allow
        adaptation to different planetary environments or mission constraints. This highlights practical
        considerations.)

        \subsection{Datasets and Input Seismogram Preparation}
            hello
            \subsubsection{Data Sources (Lunar and Martian)}
                hello
                \begin{enumerate}[label=\alph*), leftmargin=3em]
                    \item \textit{Apollo ALSEP Data:} This dataset comprises seismic recordings from the Apollo Lunar
                    Surface Experiments Package, offering crucial insights into lunar seismicity and the Moon's internal
                    structure.
                    \item \textit{Mars InSight Mission Data:} Waveforms from NASA's Mars InSight mission, specifically from
                    the SEIS instrument, provide a wealth of information on current Martian seismic activity and atmospheric
                    phenomena.
                \end{enumerate}
            \subsubsection{Waveform Segmentation and Initial Processing}
                hello
